---
title: "Population and household dynamic"
output: github_document
---

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      fig.width = 7,
                      fig.height = 5)
library(dymiumCore)
```

# Introduction

This is an tiny example of a dynamic microsimulation model for modelling population and household dynamic.

## Download the `demography` module

This example uses the `demography` module from https://github.com/dymium-org/dymiumModules.

```{r, eval = FALSE}
library(dymiumCore)
download_module(name = "demography")
```

## Set active scenario

```{r}
dymiumCore::set_active_scenario(name = "population-dynamic")
```

A friendly warning, If you feel that the log messages are too much for you can change it using the follow command.

```{r eval = FALSE}
library(lgr)
lgr::lgr$set_threshold(level = "warn")
```

To set it back to info level simply do `lgr::lgr$set_threshold(level = "info")`. But by default the log threshold is set to the 'info' level.

## Import events

```{r}
library(modules)
library(here)
event_demography_age <- modules::use(here::here("modules/demography/age.R"))
event_demography_birth <- modules::use(here::here("modules/demography/birth.R"))
event_demography_death <- modules::use(here::here("modules/demography/death.R"))
event_demography_marriage <- modules::use(here::here("modules/demography/marriage.R"))
event_demography_separation <- modules::use(here::here("modules/demography/separation.R"))
event_demography_divorce <- modules::use(here::here("modules/demography/divorce.R"))
event_demography_breakup <- modules::use(here::here("modules/demography/breakup.R"))
event_demography_cohabit <- modules::use(here::here("modules/demography/cohabit.R"))
event_demography_leavehome <- modules::use(here::here("modules/demography/leavehome.R"))
event_demography_migration <- modules::use(here::here("modules/demography/migration.R"))
```

Alternatively, the `modules` package also allow importing all R scripts from a directory, as below.

```{r eval = FALSE}
demography <- modules::use(here::here("modules/demography"))
```

Now each event can be referred by their name, eg `demography$age` instead of `event_demography_age`. This is a matter of style as both of the options work the same.


## Setup a world

```{r}
world <- World$new()

# add agents
world$add(x = Population$new(ind_data = toy_individuals, 
                             hh_data = toy_households, 
                             pid_col = c("pid", "partner_id", "father_id", "mother_id"), 
                             hid_col = "hid"))

# add models
models <- list(fertility = list(yes = 0.05, no = 0.95),
              birth_multiplicity = list("single" = 0.97, "twins" = 0.03),
              birth_sex_ratio = list(male = 0.51, female = 0.49),
              death = list(yes = 0.1, no = 0.9),
              marriage_cohab_male = list(yes = 0.1, no = 0.9),
              marriage_no_cohab_male = list(yes = 0.1, no = 0.9),
              marriage_no_cohab_female = list(yes = 0.1, no = 0.9),
              separate_male = list(yes = 0.1, no = 0.9),
              separate_child_custody = list(male = 0.2, female = 0.8),
              separate_hhtype = list(lone = 0.5, group = 0.5),
              separate_hf_random_join = list("1" = 0.4, "2" = 0.3, "3" = 0.2, "4" = 0.1),
              divorce_male = list(yes = 0.5, no = 0.9),
              divorce_female = list(yes = 0.5, no = 0.9),
              cohabitation_male = list(yes = 0.1, no = 0.9),
              cohabitation_female = list(yes = 0.1, no = 0.9),
              breakup = list(yes = 0.1, no = 0.9),
              breakup_child_custody = list(male = 0.2, female = 0.8),
              breakup_hhtype = list(lone = 0.5, group = 0.5),
              breakup_hf_random_join = list("1" = 0.4, "2" = 0.3, "3" = 0.2, "4" = 0.1),
              leavehome_male = list(yes = 0.3, no = 0.7),
              leavehome_female = list(yes = 0.2, no = 0.8),
              leavehome_hhtype = list(lone = 0.2, group = 0.8),
              leavehome_hf_random_join = list("1" = 0.5, "2" = 0.3, "3" = 0.1, "4" = 0.1),
              migrant_individuals = dymiumCore::toy_individuals, 
              migrant_households = dymiumCore::toy_households)

for (i in seq_along(models)) {
  world$add(models[[i]], name = names(models)[i])
}
```

Please see [the README page](https://github.com/dymium-org/dymiumModules/tree/master/modules/demography) of the module which serves as its documentation. 

# Setup a simulation pipeline

Let's set up a simulation pipeline using the events from the `demography` module. A for-loop is used to iteratively run the pipeline. `$start_iter` manipulates the simulation clock and pass the `world` object down to the events down the simulation pipeline.

```{r}
for (i in 1:10) {
  message("iteration: ", i)
  world$start_iter(time_step = i, unit = "year") %>%
    event_demography_age$run(.) %>%
    event_demography_birth$run(.) %>%
    event_demography_death$run(.) %>%
    event_demography_cohabit$run(.) %>%
    event_demography_marriage$run(.) %>%
    event_demography_divorce$run(.) %>%
    event_demography_breakup$run(.) %>%
    event_demography_separation$run(.) %>%
    event_demography_leavehome$run(.) %>%
    event_demography_migration$run(., target = 10)
}
```

The migration event needs target to determine the number of migrants to join the main population. Target is the number of migrant households to be added. Hence, it is unlikely to get the same number of migrant individuals in every iteration.

# Randomise the order of events

Sometimes there isn't a clear order in which a group of events should be run or you just simply want to test the order impact on your model. A group of events can be randomly shuffle in each iteration by creating a `dymiumCore::Pipeline` object and insert event functions inside. Note that, the order in which you insert the event functions in side your `Pipeline` object doesn't matter at all, since the order will be randomised.

```{r}
HouseholdFormationPipe <- Pipeline$new(
  . %>%
    event_demography_cohabit$run(.) %>%
    event_demography_marriage$run(.) %>%
    event_demography_divorce$run(.) %>%
    event_demography_breakup$run(.) %>%
    event_demography_separation$run(.)
)
```

`dymiumCore::Pipeline` has a `run(x, shuffle)` method which allows it to fit into the main simulation pipeline, as shown below. Setting the `shuffle` argument to `TRUE` will make the event functions, inside the Pipeline object, run in a random order.

```{r}
for (i in 11:20) {
  message("iteration: ", i)
  world$start_iter(time_step = i, unit = "year") %>%
    event_demography_age$run(.) %>%
    event_demography_birth$run(.) %>%
    event_demography_death$run(.) %>%
    HouseholdFormationPipe$run(., shuffle = T) %>%
    event_demography_leavehome$run(.) %>%
    event_demography_migration$run(., target = 10)
}
```

# Visualise the simulation result

```{r}
library(ggplot2)
library(patchwork)
library(scales)
history_data <- dymiumCore::get_history(world)
plot_history(x = world, by_entity = T) +
  facet_wrap(~ event) +
  guides(fill = 'none') +
  labs(y = "Occurrences", x = "Time")
```


```{r}
log_data <- dymiumCore::get_log(world)
log_data %>%
  .[grepl("^cnt:|^avl:", desc), ] %>%
  .[, value := unlist(value)] %>%
  .[, group_label := gsub(".*:", "" , x = desc)] %>% 
  .[, group_label := gsub("_", " " , x = group_label)] %>% 
  .[, tag := gsub(":.*", "" , x = desc)] %>% .[] %>%
  ggplot(data = ., aes(x = time, y = value, group = desc, color = tag)) +
  geom_line() +
  facet_wrap(~ group_label, scales = "free", labeller = labeller(group_label = label_wrap_gen(25))) +
  scale_x_continuous(breaks = scales::pretty_breaks()) +
  scale_y_continuous(breaks = scales::pretty_breaks())
```

__cnt:__ occurences

__avl:__ the total number agents that were eligible to undergo the event.
